---
title: "Nitrogen balance"
author: "Tsun-Fung Yau, Tze-Li, Manuel Händel, Eva Lorrez"
date: "28 7 2020"
output: pdf_document
---


# Questions
## 1. What are the main components of an ecosystem N budget?



## 2. What are the major N cycling processes in soils? Briefly describe the N cycle in ecosystems.

N2 -> NH4:
biological/industrial N2 fixation
NH4 -> NO3:
nitrification
NH4 <-> micr. biom.:
immobilisation
ammonification
NH4 -> NH3:
volatilization
NH4 -> Plant:
uptake
NO3 -> Plant:
uptake
NO3 leaching
NO3 -> N2:
denitrification
micr. biom <-> SOM:
depolymerisation
senescence
Plant -> SOM:
senescence

# Modelling Exercise

## 1. Run the control (con) scenarios for the 3 different management options
**rice systems (wetrice, mixedrice, dryrice), these are given on your vm. Plot
the biomass development on a daily basis from the physiology output and discuss
differences among management options.**

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)

setwd("C:/Users/Manuel/Documents/Studium/MSc_Umweltwissenschaften/model_env_sys/VM_share")

wetrice <- read.table("PH_losbanos-icon-wetrice_output/PH_losbanos-icon-wetrice-con_physiology-daily.txt",
                      header = TRUE, sep = "\t", stringsAsFactors = FALSE)

dryrice <- read.table("PH_losbanos-icon-mixedrice_output/PH_losbanos-icon-mixedrice-con_physiology-daily.txt",
                      header = TRUE, sep = "\t", stringsAsFactors = FALSE)

mixedmaize <- read.table("PH_losbanos-icon-mixedmaize_output/PH_losbanos-icon-mixedmaize-con_physiology-daily.txt",
                         header = TRUE, sep = "\t", stringsAsFactors = FALSE)

wetrice <- wetrice %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

dryrice <- dryrice %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

mixedmaize <- mixedmaize %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

w <- wetrice %>% ggplot() +
  geom_line(aes(x = datetime, y = DW_below.kgDWm.2. + DW_above.kgDWm.2.)) +
  ggtitle("Wetrice") +
  labs(x = "Date", y = "Dry Weight of Biomass [kg/m^2]")
d<- dryrice %>% ggplot() +
  geom_line(aes(x = datetime, y = DW_below.kgDWm.2. + DW_above.kgDWm.2.)) +
  ggtitle("Dryrice") +
  labs(x = "Date", y = "Dry Weight of Biomass [kg/m^2]")
m <- mixedmaize %>% ggplot() +
  geom_line(aes(x = datetime, y = DW_below.kgDWm.2. + DW_above.kgDWm.2.)) +
  ggtitle("Mixedmaize") +
  labs(x = "Date", y = "Dry Weight of Biomass [kg/m^2]")

gridExtra::grid.arrange(w,d,m)
```

The three different management options differ from year 2012 onwards. The
*dryrice* system accumulates less biomass during the dry season (first peak of
the year) due to the lack of irrigation. The *mixedmaize* system in contrast,
accumulates more biomass during the dry season, since the rice is changed for
corn crops.


## 2. Develop and compare an N-balance for the different management options
**in rice systems (wetrice, mixedrice, dryrice). Use the soilchemistry yearly
output files as well as harvest reports. Discuss your results.**


```{r, message=FALSE}
setwd("C:/Users/Manuel/Documents/Studium/MSc_Umweltwissenschaften/model_env_sys/VM_share")
soil_w <- read.table("PH_losbanos-icon-wetrice_output/PH_losbanos-icon-wetrice-con_soilchemistry-yearly.txt",
                     header = TRUE, sep = "\t", stringsAsFactors = FALSE)

soil_d <- read.table("PH_losbanos-icon-mixedrice_output/PH_losbanos-icon-mixedrice-con_soilchemistry-yearly.txt",
                     header = TRUE, sep = "\t", stringsAsFactors = FALSE)

soil_m <- read.table("PH_losbanos-icon-mixedmaize_output/PH_losbanos-icon-mixedmaize-con_soilchemistry-yearly.txt",
                     header = TRUE, sep = "\t", stringsAsFactors = FALSE)

harvest_w <- read.table("PH_losbanos-icon-wetrice_output/PH_losbanos-icon-wetrice-con_report-harvest.txt",
                        header = TRUE, sep = "\t", stringsAsFactors = FALSE)

harvest_d <- read.table("PH_losbanos-icon-mixedrice_output/PH_losbanos-icon-mixedrice-con_report-harvest.txt",
                        header = TRUE, sep = "\t", stringsAsFactors = FALSE)

harvest_m <- read.table("PH_losbanos-icon-mixedmaize_output/PH_losbanos-icon-mixedmaize-con_report-harvest.txt",
                        header = TRUE, sep = "\t", stringsAsFactors = FALSE)



soil_w <- soil_w %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

soil_d <- soil_d %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

soil_m <- soil_m %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

harvest_w <- harvest_w %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

harvest_d <- harvest_d %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))

harvest_m <- harvest_m %>% 
  mutate_at("datetime", lubridate::ymd_hms) %>% 
  mutate(year = year(datetime))


n_cols <- grep("aN", names(soil_w), ignore.case = FALSE, value = TRUE)
n_cols <- grep("litter|mineral|nitrify|chemo|uptake|change", n_cols, value = TRUE, invert = TRUE)


soil_w <- select(soil_w, c(n_cols, ncol(soil_w))) %>%
  rename_at(1:length(n_cols), function(name) sub(pattern = ".*?_(.*)\\..*\\..*\\..*", replacement = "\\1", name))
  
soil_d <- select(soil_d, c(n_cols, ncol(soil_d))) %>%
  rename_at(1:length(n_cols), function(name) sub(pattern = ".*?_(.*)\\..*\\..*\\..*", replacement = "\\1", name))

soil_m <- select(soil_m, c(n_cols, ncol(soil_m))) %>%
  rename_at(1:length(n_cols), function(name) sub(pattern = ".*?_(.*)\\..*\\..*\\..*", replacement = "\\1", name))


harvest_w <- harvest_w %>% select(c(13,15,18)) %>% 
  transmute(year = year,
            n_harvest = dN_bud_export.kgNha.1. + dN_straw_export.kgNha.1.) %>% 
  group_by(year) %>% 
  summarise(n_harvest = sum(n_harvest))

harvest_d <- harvest_d %>% select(c(13,15,18)) %>% 
  transmute(year = year,
            n_harvest = dN_bud_export.kgNha.1. + dN_straw_export.kgNha.1.) %>% 
  group_by(year) %>% 
  summarise(n_harvest = sum(n_harvest))

harvest_m <- harvest_m %>% select(c(13,15,18)) %>% 
  transmute(year = year,
            n_harvest = dN_bud_export.kgNha.1. + dN_straw_export.kgNha.1.) %>% 
  group_by(year) %>% 
  summarise(n_harvest = sum(n_harvest))

balance_w <- inner_join(harvest_w, soil_w, by = "year") %>% 
  mutate_at(2:9, function(value) -value) %>% 
  mutate(balance = rowSums(.[2:13])) %>% 
  filter(year %in% 2012:2017) %>% 
  summarise_all(mean) %>% 
  mutate(year = "wetrice")%>%
  rename(system =  year)

balance_d <- inner_join(harvest_d, soil_d, by = "year") %>% 
  mutate_at(2:9, function(value) -value) %>% 
  mutate(balance = rowSums(.[2:13])) %>% 
  filter(year %in% 2012:2017) %>% 
  summarise_all(mean) %>% 
  mutate(year = "dryrice") %>%
  rename(system =  year)

balance_m <- inner_join(harvest_m, soil_m, by = "year") %>% 
  mutate_at(2:9, function(value) -value) %>% 
  mutate(balance = rowSums(.[2:13])) %>% 
  filter(year %in% 2012:2017) %>% 
  summarise_all(mean) %>% 
  mutate(year = "mixedmaize")%>%
  rename(system =  year)

balance <- rbind(balance_w, balance_d, balance_m)

balance %>% pivot_longer(2:14) %>%
  ggplot() +
  geom_bar(aes(x = name, y = value, fill = name), stat = "identity") +
  labs(x = "", y = "Nitrogen flux [kgN/ha]") +
  theme(axis.text.x = element_blank()) +
  # theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  facet_wrap(vars(system))
```

The values displayed are the average values from 2012 to 2017. The ovearall
balance of the 3 systems is always negative. N2 and N2O emissions are lower when
the crops are being irrigated during the dry season (see wetrice). The
*mixedmaize* system provides a bigger harvest.

## 3. Assess the change of fertilizer type on the N-balance for one of these management systems. Plot the results
on the N budget and discuss the results. For this, make changes in the event files and simulate the following
fertilizer types:
• anhydrous ammonia: "nh3“  
• nitrate: "no3”  
• urea: "urea“  
• ammonium nitrate: "nh4no3” since it is frequently used  
```{r}

```

