---
title: "Exercise 2"
author: "Manuel Händel"
date: "20 7 2020"
output: pdf_document
---

# 1. Install and load the R package ncdf4

```{r}
library(ncdf4)
```

# 2. Read in the netcdf and explore the structure of the data.

```{r}
prcp <- nc_open(file.path(getwd(), "day01/material/prcp_daily_2000-2000.nc"))
tmax <- nc_open(file.path(getwd(), "day01/material/tmax_daily_2000-2000.nc"))
tmin <- nc_open(file.path(getwd(), "day01/material/tmin_daily_2000-2000.nc"))

str(tmin)

```


# 3. Plot a world map of maximum temperature at one selected day and check the orientation of the data.
# 4. Select now a subset of the data which is the time-series at the grid-cell containing Freiburg.
Plot the time
series of precipitation and min/max temperature. Make sure axes and legend are labelled meaningfully with
the correct unit. What was the maximum temperature in the year 2000?
# 5. Calculate the average temperature from the minimum and maximum temperatures.
# 6. Aggregate the daily data to monthly values.
Use the adequate way of aggregating for the respective variable
(e.g. mean or sum). Create a data.frame with monthly values and a header containing Year, Month, Temp
(temperature in degree C), Precip (precipitation in mm). Check the annual precipitation sum and look up if
they make sense.
# 7. The data.frame contains a single year only. 
Search for climate projections for two different RCPs and
construct two artificial climate datasets of future scenarios covering 100 more years.
# 8. Construct a further artificial climate dataset
where of future climate scenario covering 100 more years is just
containing repeatedly the temperature and precipitation data of the year 2000 (reference scenario).
# 9. Store your datasets as tab-seperated .txt files
using meaningful filenames, you will need them for another
exercise.



